---
title: "Integrating Python and R"
output: html_document
author: Esteban López Ochoa
inst: Master in Science in Urban Planning, UTSA
---

# Python in R Markdown

In R many things can be done, but there are some functionalities that python is just better at. Another reason to include `python` whinin `R`, is that some users might know to do (or be more comfortable doing) something in `python`, but other things in `R`. This script provides a short introduction on how to run python within R-Studio.

For example, bellow there is some R code to produce a graph within R.

```{r `Example 1: Data in R`}
data("LifeCycleSavings")
summary(LifeCycleSavings)
?LifeCycleSavings
head(LifeCycleSavings)
#install.packages('ggplot2')
library(ggplot2)
typeof(LifeCycleSavings); class(LifeCycleSavings)
LifeCycleSavings$rn<-row.names(LifeCycleSavings)
ggplot(LifeCycleSavings)+geom_bar(aes(x = reorder(rn,sr),y=sr),stat = 'identity')+coord_flip()
```

However there might be times that is desirable to bring some python functionalities to your `R` workflow. In this cases, there are several ways to do so.

# 1. Python in R-Studio and RMarkdown using the `reticulate` package

The package `reticulate` provides an easier way to integrate python in R. The basic set-up process includes.
- Installing and loading `reticulate`
- Checking if python is installed and what version this is
- Checking the virual and conda environments that are available
- Checking if the Markdown engine points to the right python version and/or environment
- Installing python packages

```{r py_setup}
#install.packages('reticulate')
library(reticulate)
repl_python() #turns your R console into a Python Console. If no python is found choose yes to install miniconda

py_discover_config() # info about your python distribution
conda_list()
virtualenv_list()
py_install('pandas') # installs python libraries 
py_install('matplotlib') # installs python libraries 
```

### Running python within a `chunk`

The following code shows how python can be runned in a R-Studio `chunk`, in order to do something (in this case a basic use of the pandas and numpy packages that for the sake of the example we will assume it is easier to do in python).

```{python}
16**0.5
import math
math.sqrt(16)
import pandas as pd
import numpy as np
LC = pd.DataFrame(r.LifeCycleSavings)
print(r.LifeCycleSavings.head())
print(LC.head())
LC.dtypes
LC.describe()
LC2 = LC.loc[:,["sr","dpi"]]
LC2.loc[LC2.loc[:,"sr"]>10,"sr"]
```


### Using python products back in R

Finnaly, once all python code has been runned, you can easily bring that output back to R, and visualize it as you like it. 

```{r back to R, echo=TRUE}
library(ggplot2)
ggplot(py$LC2,aes(x=reorder(row.names(py$LC2),dpi),y=dpi))+geom_bar(stat = 'identity')+coord_flip()
```

# 2. Turning the R-Console as a python-Console

The function `repl_python()` is a wrapper that allows turning your R-Console as a python console. This will allow you to run `python` code creating objects in that environment that can later be returned to your `R` environment. For more information visit: https://rstudio.github.io/reticulate/ 

# 3. Sourcing an entire python script in R

Python scripts can have either code processing data, and other programming activities, but can also have functions. For this latte type of needs, sourcing a python script can be quite useful. This can be done by using the function `source_python` in the `reticulate` package. For more information visit: https://rstudio.github.io/reticulate/ 

# 4. Running R within your python distribution 

If you have installed the Anaconda suite. There are two ways in which you can bring R to your python work flow.

## A. Jupyter Notebooks

The `kernel` of a Jypyter Notebook can be switched from the dafault `python` engine to an `R` engine. This can be done by creating an `R` based environment in your python and then, using that environment as the active environment, launching a Jupyter Notebook. For more information visit:
https://docs.anaconda.com/anaconda/navigator/tutorials/create-r-environment/ 

## B. R-Studio in the Anaconda suite.

This is just a link for opening R-Studio from the Anaconda Navigator.
Visit: https://docs.anaconda.com/anaconda/navigator/tutorials/create-r-environment/ 
